Intersection COVID-19: Italy

### 2020-07-01 README.Rmd
fixed fips numeric

[![](https://img.shields.io/badge/devel%20version-0.0.0.9800-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
[![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)
[![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)

>>>>

[![](https://img.shields.io/badge/devel%20version-0.0.0.9900-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
[![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)
[![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)

### 2020-07-01 notebooks/old_notebooks/report_italy_wp.Rmd
fixed fips numeric

@@

```{r dem_it_p}
dem_it_p <- vroom(
  "data/italy/pop_it.csv",
  col_types = cols(
    ITTER107 = col_character(),
    Territorio = col_character(),
    TIPO_DATO15 = col_character(),
    `Tipo di indicatore demografico` = col_character(),
    SEXISTAT1 = col_double(),
    Sesso = col_character(),
    ETA1 = col_character(),
    Età  = col_character(),
    STATCIV2 = col_double(),
    `Stato civile` = col_character(),
    TIME = col_double(),
    `Seleziona periodo` = col_double(),
    Value = col_double(),
    `Flag Codes` = col_character(),
    Flags = col_character()
  )
) %>%
  rename(
    region = Territorio,
    index = `Tipo di indicatore demografico`,
    sex = Sesso,
    age = `Età`,
    marital_status = `Stato civile`,
    year = `Seleziona periodo`,
    value = Value
  ) %>%
  select(region, index, sex, age, marital_status, year, value) %>%
  filter(age != "totale", sex != "totale", marital_status == "totale") %>%
  mutate(
    age = as.numeric(str_remove_all(age, "[[:alpha:]]")),
    sex = recode(sex, femmine = "female", maschi = "male")
  )

@@

chronic_it_p <- vroom("data/italy/chronic_conditions.csv",
  col_types = cols(
    ITTER107 = col_character(),
    Territorio = col_character(),
    TIPO_DATO_AVQ = col_character(),
    `Tipo dato` = col_character(),
    MISURA_AVQ = col_character(),
    Misura = col_character(),
    TIME = col_double(),
    `Seleziona periodo` = col_double(),
    Value = col_double(),
    `Flag Codes` = col_logical(),
    Flags = col_logical()
  )
) %>%
  clean_names() %>%
  filter(misura == "valori in migliaia") %>% # we keep the absolute values
  mutate(value = value * 1000) # values are in thousands

@@

```{r smoking_it, warning=FALSE, message=FALSE}
smoking_it <- vroom("data/italy/smoking.csv",
  col_types = cols(
    ITTER107 = col_character(),
    Territorio = col_character(),
    TIPO_DATO_AVQ = col_character(),
    `Tipo dato` = col_character(),
    MISURA_AVQ = col_character(),
    Misura = col_character(),
    TIME = col_double(),
    `Seleziona periodo` = col_double(),
    Value = col_double(),
    `Flag Codes` = col_logical(),
    Flags = col_logical()
  )
) %>%
  rename(
    region = Territorio,
    smoking_status = TIPO_DATO_AVQ,
    value = Value
  ) %>%
  filter(Misura == "valori in migliaia") %>%
  select(region, smoking_status, value) %>%
  mutate(region = str_replace(region, "-", " ")) %>%
  filter(!region %in% c("Trentino Alto Adige / Südtirol", "Italia")) %>%
  # make name consistent
  mutate(region = recode(region,
    "Valle d'Aosta / Vallée d'Aoste" = "Valle d'Aosta",
    "Provincia Autonoma Bolzano / Bozen" =  "P.A. Bolzano",
    "Provincia Autonoma Trento"   =  "P.A. Trento",
    "Fiuli Venezia Giulia"  = "Friuli Venezia Giulia"
  )) %>%
  mutate(smoking_status = recode(smoking_status,
    "14_FUMO_SI" = "smoking_current",
    "14_FUMO_EX" = "smoking_ex",
    "14_FUMO_NO" = "smoking_no"
  )) %>%
  filter(smoking_status %in% c("smoking_current", "smoking_ex", "smoking_no")) %>%
  # values are in thousands
  mutate(value = value * 1000) %>%
  pivot_wider(names_from = smoking_status, values_from = value)
```

@@

```{r obesity_it, warning=FALSE, message=FALSE}
bweight_it <- vroom("data/italy/bweight_it.csv", col_types = cols(
  ITTER107 = col_character(),
  Territorio = col_character(),
  TIPO_DATO_AVQ = col_character(),
  `Tipo dato` = col_character(),
  MISURA_AVQ = col_character(),
  Misura = col_character(),
  SEXISTAT1 = col_double(),
  Sesso = col_character(),
  TIME = col_double(),
  `Seleziona periodo` = col_double(),
  Value = col_double(),
  `Flag Codes` = col_logical(),
  Flags = col_logical()
)) %>%
  filter(TIME == 2018) %>%
  rename(
    region = Territorio,
    bweight_status = TIPO_DATO_AVQ,
    value = Value,
    ses = Sesso
  ) %>%
  filter(Misura == "valori in migliaia") %>%
  select(region, bweight_status, value) %>%
  mutate(region = str_replace(region, "-", " ")) %>%
  filter(!region %in% c("Trentino Alto Adige / Südtirol", "Italia")) %>%
  # make name consistent
  mutate(region = recode(region,
    "Valle d'Aosta / Vallée d'Aoste" = "Valle d'Aosta",
    "Provincia Autonoma Bolzano / Bozen" =  "P.A. Bolzano",
    "Provincia Autonoma Trento"   =  "P.A. Trento",
    "Fiuli Venezia Giulia"  = "Friuli Venezia Giulia"
  )) %>%
  mutate(bweight_status = recode(bweight_status,
    "18_BMI_SOTTO" = "bweight_under",
    "18_BMI_NORMO" = "bweight_normal",
    "18_BMI_SOVRA" = "bweight_over",
    "18_BMI_OBE" = "bweight_obese"
  )) %>%
  filter(bweight_status %in% c("bweight_under", "bweight_normal", "bweight_over", "bweight_obese")) %>%
  # values are in thousands
  mutate(value = value * 1000) %>%
  pivot_wider(names_from = bweight_status, values_from = value)
```

@@

@@

>>>>

# Objective

Investigate the relation (if any) between influenza vaccination coverage and measures related to COVID-19 epidemic in Italy. Demigrafical and health related data of each italian region are also taken into account.

@@

```{r dem_it_p}
dem_it_p <- vroom(
  "/Users/heverz/Documents/R_projects/covid19_interference/data/italy/pop_it.csv",
  col_types = cols(
    ITTER107 = col_character(),
    Territorio = col_character(),
    TIPO_DATO15 = col_character(),
    `Tipo di indicatore demografico` = col_character(),
    SEXISTAT1 = col_double(),
    Sesso = col_character(),
    ETA1 = col_character(),
    Età  = col_character(),
    STATCIV2 = col_double(),
    `Stato civile` = col_character(),
    TIME = col_double(),
    `Seleziona periodo` = col_double(),
    Value = col_double(),
    `Flag Codes` = col_character(),
    Flags = col_character()
  )
) %>%
  rename(
    region = Territorio,
    index = `Tipo di indicatore demografico`,
    sex = Sesso,
    age = `Età`,
    marital_status = `Stato civile`,
    year = `Seleziona periodo`,
    value = Value
  ) %>%
  select(region, index, sex, age, marital_status, year, value) %>%
  filter(age != "totale", sex != "totale", marital_status == "totale") %>%
  mutate(
    age = as.numeric(str_remove_all(age, "[[:alpha:]]")),
    sex = recode(sex, femmine = "female", maschi = "male")
  )

@@

chronic_it_p <- vroom("/Users/heverz/Documents/R_projects/covid19_interference/data/italy/chronic_conditions.csv",
  col_types = cols(
    ITTER107 = col_character(),
    Territorio = col_character(),
    TIPO_DATO_AVQ = col_character(),
    `Tipo dato` = col_character(),
    MISURA_AVQ = col_character(),
    Misura = col_character(),
    TIME = col_double(),
    `Seleziona periodo` = col_double(),
    Value = col_double(),
    `Flag Codes` = col_logical(),
    Flags = col_logical()
  )
) %>%
  clean_names() %>%
  filter(misura == "valori in migliaia") %>% # we keep the absolute values
  mutate(value = value * 1000) # values are in thousands

@@

```{r smoking_it, warning=FALSE, message=FALSE}
smoking_it <- vroom("/Users/heverz/Documents/R_projects/covid19_interference/data/italy/smoking.csv",
  col_types = cols(
    ITTER107 = col_character(),
    Territorio = col_character(),
    TIPO_DATO_AVQ = col_character(),
    `Tipo dato` = col_character(),
    MISURA_AVQ = col_character(),
    Misura = col_character(),
    TIME = col_double(),
    `Seleziona periodo` = col_double(),
    Value = col_double(),
    `Flag Codes` = col_logical(),
    Flags = col_logical()
  )
) %>%
  rename(
    region = Territorio,
    smoking_status = TIPO_DATO_AVQ,
    value = Value
  ) %>%
  filter(Misura == "valori in migliaia") %>%
  select(region, smoking_status, value) %>%
  mutate(region = str_replace(region, "-", " ")) %>%
  filter(!region %in% c("Trentino Alto Adige / Südtirol", "Italia")) %>%
  # make name consistent
  mutate(region = recode(region,
    "Valle d'Aosta / Vallée d'Aoste" = "Valle d'Aosta",
    "Provincia Autonoma Bolzano / Bozen" =  "P.A. Bolzano",
    "Provincia Autonoma Trento"   =  "P.A. Trento",
    "Fiuli Venezia Giulia"  = "Friuli Venezia Giulia"
  )) %>%
  mutate(smoking_status = recode(smoking_status,
    "14_FUMO_SI" = "smoking_current",
    "14_FUMO_EX" = "smoking_ex",
    "14_FUMO_NO" = "smoking_no"
  )) %>%
  filter(smoking_status %in% c("smoking_current", "smoking_ex", "smoking_no")) %>%
  # values are in thousands
  mutate(value = value * 1000) %>%
  pivot_wider(names_from = smoking_status, values_from = value)
```

@@

```{r obesity_it, warning=FALSE, message=FALSE}
bweight_it <- vroom("/Users/heverz/Documents/R_projects/covid19_interference/data/italy/bweight_it.csv", col_types = cols(
  ITTER107 = col_character(),
  Territorio = col_character(),
  TIPO_DATO_AVQ = col_character(),
  `Tipo dato` = col_character(),
  MISURA_AVQ = col_character(),
  Misura = col_character(),
  SEXISTAT1 = col_double(),
  Sesso = col_character(),
  TIME = col_double(),
  `Seleziona periodo` = col_double(),
  Value = col_double(),
  `Flag Codes` = col_logical(),
  Flags = col_logical()
)) %>%
  filter(TIME == 2018) %>%
  rename(
    region = Territorio,
    bweight_status = TIPO_DATO_AVQ,
    value = Value,
    ses = Sesso
  ) %>%
  filter(Misura == "valori in migliaia") %>%
  select(region, bweight_status, value) %>%
  mutate(region = str_replace(region, "-", " ")) %>%
  filter(!region %in% c("Trentino Alto Adige / Südtirol", "Italia")) %>%
  # make name consistent
  mutate(region = recode(region,
    "Valle d'Aosta / Vallée d'Aoste" = "Valle d'Aosta",
    "Provincia Autonoma Bolzano / Bozen" =  "P.A. Bolzano",
    "Provincia Autonoma Trento"   =  "P.A. Trento",
    "Fiuli Venezia Giulia"  = "Friuli Venezia Giulia"
  )) %>%
  mutate(bweight_status = recode(bweight_status,
    "18_BMI_SOTTO" = "bweight_under",
    "18_BMI_NORMO" = "bweight_normal",
    "18_BMI_SOVRA" = "bweight_over",
    "18_BMI_OBE" = "bweight_obese"
  )) %>%
  filter(bweight_status %in% c("bweight_under", "bweight_normal", "bweight_over", "bweight_obese")) %>%
  # values are in thousands
  mutate(value = value * 1000) %>%
  pivot_wider(names_from = bweight_status, values_from = value)
```

@@

With this we check that the regions are the same across datasets. They are not always very consistent with names.
```{r check_join_all}
# check that the names are right otherwise when we join we loose data
# lapply(list(
#   cmr_it,
#   dem_it_wider,
#   regions_area,
#   chronic_it,
#   cancer_it,
#   smoking_it,
#   bweight_it
# ), function(x) {
#   unique(x$region) %in% unique(dem_it_wider$region)
# })
```

@@

## Now We Rock`n`roll baby

Let' start looking at COVID-19 measures.

```{r}
unique(all_dat$COVID19_var)
```

### total_cases

```{r total_cases, fig.height= 6, fig.width=6}
all_dat %>% 
    filter(COVID19_var == "total_cases") %>% 
    ggplot(aes(date, value, group = region, color = region)) + 
      geom_line() +
      # geom_smooth(method = "lm", formula = y~x, se = FALSE) +
      scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
      theme_bw() +
      theme(legend.position = "none") +
      facet_wrap(vars(region)) +
        labs(
        x = "date",
        y = "confirmed cases") 
```
```{r total_cases_log, fig.height= 6, fig.width=6}
all_dat %>% 
    filter(COVID19_var == "total_cases") %>% 
    ggplot(aes(date, log2(value), group = region, color = region)) + 
      geom_line() +
      # geom_smooth(method = "lm", formula = y~x, se = FALSE) +
      scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
      theme_bw() +
      theme(legend.position = "none") +
      facet_wrap(vars(region)) +
        labs(
        x = "date",
        y = "confirmed cases (log2)") 
```
Clearly at the extreme of date we lose linearity. But that might make sense

Let's do some smooth analysis.

```{r total_cases_smooth, fig.height= 6, fig.width=6, warning=FALSE}
p <- ggplotly(all_dat %>% 
    filter(COVID19_var == "total_cases") %>% 
    ggplot(aes(date, log2(value), group = region, color = region)) + 
      # geom_line() +
      geom_smooth(method = "lm", formula = y~x, se = FALSE, lwd = 0.5) +
      scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
      theme_bw() +
      theme(legend.position = "none") +
      # facet_wrap(vars(region)) +
        labs(
        x = "date",
        y = "confirmed cases"),
    tooltip = c("date", "log2(value)", "group")

    ) %>% 
   config(p = ., displayModeBar = FALSE)

p
```

Well beside `Piemonte` and `Molise`, rate is very similar.

othe vars:

cmr: altready linear. saturation hospitals
chr: funny at the beginning they were detecting only peple that were hospitalizing
cir:
```{r mortality_log, fig.height= 6, fig.width=6, message=FALSE, warning=FALSE}
all_dat %>% 
    filter(COVID19_var == "chr") %>% 
    ggplot(aes(date, value, group = region, color = region)) + 
      geom_line() +
      # geom_smooth(method = "lm", formula = y~x, se = FALSE) +
      scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
      theme_bw() +
      theme(legend.position = "none") +
      facet_wrap(vars(region)) +
        labs(
        x = "date")
```

<!-- This chunk below save datasets and data in my folder that then sync with dropbox. -->
```{r savedata, include=FALSE, eval=FALSE}
# files   ############
# report_workingprogress: notebook with all the progresses
# renv.lock: file to restore libraries
saveRDS(all_dat, "data/italy/all_dat.RDS")

drop_upload("data/italy/all_dat.RDS", path = "covid19/interference/data/")
drop_upload("tidy_data_script.R", path = "covid19/interference/code/")
drop_upload("data/italy/all_dat.RDS", path = "covid19/interference/data/italy/")
drop_upload("report_italy_wp.html", path = "covid19/interference/text/")
drop_upload("renv.lock", path = "covid19/interference/renv/")
drop_upload("report_italy_wp.Rmd", path = "covid19/interference/code/")
drop_upload("data/italy/NPI_it.xlsx", path = "covid19/interference/data/italy/raw/")
```
